---
title: "Analysing Helipad Data"
author: "Danny Wong"
date: "Sunday, April 05, 2015"
output: html_document
---

We have some data from the KSS HEMS patients conveyed to KCH.

```{r, message=FALSE}
#Read the data into a dataframe, specify the NA strings to incorporate the missing values
data <- read.csv("data/HEMSdata.csv", na.strings=c("NA","n/a", ""))

#Let's look at the table
head(data)

#Let's see how many cases there are
print(nrow(data))
```

Notice that the coordinates are in the Ordnance Survey Grid format under the column "Grid". For us to work with this meaningfully, we need to convert this to latitude and longitude (WGS84) coordinates.

Fortunately there are others who have encountered the same problem before:

Sources: 

    * http://stackoverflow.com/questions/23017053/how-to-convert-uk-grid-reference-to-latitude-and-longitude-in-r
    * https://stat.ethz.ch/pipermail/r-sig-geo/2010-November/010141.html
    * http://www.hannahfry.co.uk/blog/2012/02/01/converting-british-national-grid-to-latitude-and-longitude-ii
    * http://cran.r-project.org/web/packages/rnrfa/rnrfa.pdf

```{r, message=FALSE}
#Load the required packages
require(rnrfa)
require(dplyr)

#Remove the rows with missing Grid references
data <- filter(data, !is.na(Grid))
  
#Parse the OS Grid References into Eastings and Northings, then pipe it into WSG84 Coordinates
coordinates <- OSGParse(data$Grid) %>% OSG2LatLon()

#Add these onto the data
data <- mutate(data, lat = coordinates$Latitude, lon = coordinates$Longitude)

#Write to .csv
write.csv(data, file="data/cleaned.csv")
```

We can now start making some maps!

Let's first visualise the locations of each HEMS pickup as a quick visualisation

```{r, message=FALSE}
#Load the required libraries
require(ggplot2)
require(ggmap)
require(dplyr)

#Get map from Google
ukmap <- get_map(location = "London, UK", zoom = 8, scale = 4)
ggmap(ukmap) + geom_point(data = data, aes(x = data$lon, y = data$lat, fill = "red", alpha = 0.8), size = 4, shape = 21) + guides(fill=FALSE, alpha=FALSE, size=FALSE)

#We can separate the cases by the county where they come from, let's see the counties where they come from
county <- count(data, County)
county

#Let's measure the distances between all the HEMS pickup sites and KCH
#We need the coordinates for KCH
KCH <- geocode("King's College Hospital, Denmark Hill, London")

#Let's load the required library to calculate distances
require(geosphere)

#We apply the VincentyEllipsoid method of calculating straight-line distance, and 
coordinates <- data.frame("lon" = coordinates$Longitude, "lat" = coordinates$Latitude)
KCH <- data.frame("lon" = KCH$lon, "lat" = KCH$lat) 
coordinates$distance <- distVincentyEllipsoid(p1 = coordinates, p2 = KCH) #This calculates distances in meters

#Lets add this distance to the master data frame, but convert it to kilometers first
data <- mutate(data, distance = coordinates$distance/1000)

summary(data$distance)

#Plot a histogram of the distances
hist(data$distance, breaks = 25, prob=TRUE, col="grey", xlab = "Distance of transfers (km)", ylab = "Relative Frequency", main = "Histogram of Transfer Distances with Kernel Density Curve")

#Overlay the kernel density plot of the distances
lines(density(data$distance))
```

We will create a new field for drawing lines between the Head and Tail sites on QGIS in a Well-Known-Text (WKT) LINESTRING format

``` {r}
#Text formatting for WKT LINESTRING
data$wkt <- paste0("LINESTRING(", data$lon, " ", data$lat, ",", KCH$lon, " ", KCH$lat, ")")

#Save the dataframe as a .csv
write.csv(data, file = "data/cleaned.csv")
write.csv(KCH, "data/KCH.csv")
```

We want to see whether there is a temporal relationship with HEMS cases, and whether there St George's building their Helipad has had any effect on KCH's workload.

```{r, message=FALSE}
#Load the required libraries
require(lubridate)
require(dplyr)

#We use the Lubridate package to manipulate the dates and time strings into POSIXct format so we can perform stats on them
timestamp <- paste(data$Date, " ", data$Time)

data$Date <- dmy(data$Date)
data$Day <- wday(data$Date, label = TRUE)

#Let's see if there is a difference between the cases depending of which day of the week they arrive
count(data, Day)

#Let's select the cases 6 months before St George's Helipad was constructed and 6 months after
data_before <- data %>% filter(Date >= dmy("01/10/2013") & Date < dmy("01/04/2014"))
data_after <- data %>% filter(Date >= dmy("01/04/2014") & Date < dmy("01/10/2014")) 

count(data_before, County)
count(data_after, County)

#Let's draw a time series of the cases, by monthly frequency
case_count <- count(data, Date)
case_count$Month <- as.Date(cut(case_count$Date, breaks = "month"))
monthly_count <- count(case_count, Month)

#Load ggplot2 to graph
require(ggplot2)

#Plot
ggplot(monthly_count, aes(Month, n)) + 
    geom_line() + 
    xlab("") + 
    ylab("Cases") + 
    ggtitle("Monthly Frequency of KSS HEMS Cases brought to KCH")

#Let's now draw a time series of the cases, by weekly frequency this time
case_count$Week <- as.Date(cut(case_count$Date, breaks = "week"))
weekly_count <- count(case_count, Week)

#Plot
ggplot(weekly_count, aes(Week, n)) + 
    geom_line() + 
    xlab("") + 
    ylab("Cases") +
    ggtitle("Weekly Frequency of KSS HEMS Cases brought to KCH")
```

